The evolution of gene expression in the term placenta of viviparous mammals

### 2015-10-19 manuscript/manuscript.Rnw
include all tissue go analysis

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","palcenta_go_results_mf")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf"
                              ))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low="white",
                            high="blue",
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

<<go.analysis,warning=FALSE,message=FALSE,echo=FALSE,results="asis">>=
go.results <- rbindlist(list(placenta_go_results_mf[1:10,],
                             placenta_go_results_bp[1:10,]))
go.results <- go.results[,Annotated:=NULL]
setnames(go.results,"classic","p value")
setnames(go.results,"classic.fdr","FDR")
setnames(go.results,"GO.ID","GO ID")
setnames(go.results,"Term","GO Term")
print(xtable(go.results,
             digits=2,
             display=c("d","s","s","d","f","e","e","s"),
             caption="Top go MF results",
             label="tab:core_placenta_go_terms"),
      include.rownames=FALSE,
      floating=TRUE,
      booktabs=TRUE,
      size="\\tiny"
      )
@

@@

>>>>

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf"
                              ))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low="white",
                            high="blue",
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

<<go.analysis,warning=FALSE,message=FALSE,echo=FALSE,results="asis">>=
placenta.go.results <- rbindlist(list(placenta_go_results_mf[1:10,],
                             placenta_go_results_bp[1:10,]))
placenta.go.results <- placenta.go.results[,Annotated:=NULL]
setnames(placenta.go.results,"classic","p value")
setnames(placenta.go.results,"classic.fdr","FDR")
setnames(placenta.go.results,"GO.ID","GO ID")
setnames(placenta.go.results,"Term","GO Term")
print(xtable(placenta.go.results,
             digits=2,
             display=c("d","s","s","d","f","e","e","s"),
             caption="Top go MF results",
             label="tab:core_placenta_go_terms"),
      include.rownames=FALSE,
      floating=TRUE,
      booktabs=TRUE,
      size="\\tiny"
      )
@

@@

\begin{table}
<<go.analysis.all.tissue,echo=FALSE,results="asis">>=
go.analysis.wide <- 
    data.table(dcast(data=go.results,
                     formula=GO.ID+type+Term+Ontology~species,
                     value.var="FDR"))
go.analysis.wide <- 
    go.analysis.wide[,`max(FDR)`:=apply(go.analysis.wide[,-(1:4),with=FALSE],
                                    1,max,na.rm=TRUE)]
go.analysis.wide.table <- 
    go.analysis.wide[`max(FDR)` <= 0.1 & grepl("No Housekeeping",type),
                     list(GO.ID,Term,Ontology,`max(FDR)`)]
go.analysis.wide.table <- 
    go.analysis.wide.table[order(`max(FDR)`)]
print(xtable(go.analysis.wide.table,
             display=c("d","s","s","s","e")
             ),
      booktabs=TRUE,
      floating=FALSE
      )
@ 
\caption{GO terms with enrichment in all species analyzed with
  housekeeping genes removed}
\label{tab:go_terms_all_species}
\end{table}
